%!TEX root=cs112S2015-lab3.tex 
% mainfile: cs112S2015-lab3.tex 

\input{labspre.tex}

\usepackage[compact]{titlesec}

\begin{document} \MYTITLE{Laboratory Assignment Three: A First Step Towards Implementing a Twitter Client}
\MYHEADERS{Laboratory Assignment Three}{Due: February 5, 2015}

\section*{Introduction}

  Twitter is a commonly used social media network.  As we refamiliarize ourselves with the fundamental characteristics
  of the object-oriented programming paradigm and continue to learn more advanced features of the Vim integrated
  development environment (IDE), we will take the first steps towards implementing a complete Twitter client. Along with
  learning how to use, enhance, and extend a Java class, we will also review the steps that you must take to perform
  console input and output. This laboratory assignment represents the first of a two-part series of assignments. 

\section*{Accessing and Understanding GatorTweet}

  To start this laboratory assignment, you should return to the {\tt cs112S2014-share} Git repository and type the
  command {\tt git pull} in the terminal window.  Now, you should have a {\tt lab3/GatorTweet/} directory that you can
  explore further.  Where is the Java source code in this new directory? Why is it stored in the directory where you
  found it? Now, you should use GVim to study the source code in the {\tt build.xml} file.  What do you think is the
  purpose of this file?  For this assignment, you can type the command ``{\tt ant compile}'' in your terminal window and
  it will compile the {\tt Tweet.java} file and save the bytecode in the correct subdirectories inside of the {\tt bin/}
  directory. 

  At this point, you should purposefully insert an error into the {\tt Tweet.java} file.  For instance, you can open the
  file in GVim and then remove one of the semi-colons at the end of a source code line. Now, type ``{\tt ant compile}''
  again.  What output do you see on the screen? Now, you can see the output of the build command will indicate where the
  error is located in your program. After you have returned to the error location, go ahead and fix the problem by
  adding back the semi-colon and recompiling your program. If you would like to further experiment with automatically
  building your Java program, you can type the command ``{\tt :Ant compile}'' directly in GVim. What action does this
  command perform? How is this approach different than running {\tt ant} in the terminal?

  It is also worth noting that the files and the directories for this assignment are organized in a way that is
  different from what you have seen in past laboratories or in-class assignments in this class or in your previous
  introductory computer science class. One difference that you should notice is that the bytecode files (i.e., the files
    ending with the ``{\tt .class} extension) are stored in a different directory that the source code files (i.e., the
  files that end with ``{\tt .java}''). What are the advantages and disadvantages of organizing the project in this
  fashion? Additionally, you should note that the source code for this project is stored in nested directories and the
  source code file called {\tt Tweet.java} contains the ``{\tt package edu.allegheny.gatortweet;}'' at the top of it.
  Why do you think that the source code is organized in this fashion? What are the benefits and drawbacks?

\section*{Improving the GatorTweet System}

  You will notice that the {\tt Tweet} class contains two instance variables and several methods.  First, you should add
  proper JavaDoc comments to all of the variables and methods in {\tt Tweet.java}.  Once you have finished adding in all
  of the necessary comments, please notice that the {\tt isValidMessage} method is not correctly implemented.  You
  should enhance this method so that it returns {\tt true} when a candidate message is valid and {\tt false} otherwise.
  This method should classify a message as valid as long as its length is greater than zero and less than or equal to
  140.

% \section*{Adding a GatorTweet Class}

  \noindent
  Now, you should implement a {\tt GatorTweet} class that performs the following operations in {\tt main}:

\vspace*{-.1in} \begin{enumerate}
    % \itemsep 0in
    \item Constructs an {\tt ArrayList} that will only hold {\tt Tweet} objects for the valid {\tt Tweet}s.

    \item Constructs an {\tt ArrayList} that will only hold {\tt String}s for the invalid tweet messages.

    \item Uses the {\tt java.util.Scanner} class to read in consecutive lines of text from a file in the {\tt tweets/}
      subdirectory of the {\tt GatorTweet/} directory, called ``Tweets.txt'', that contains potential tweets. (It is
      worth noting that this file can contain both valid and invalid tweets).

    \item Checks to see if the current tweet message is a valid one or not.  If the message is valid, then it should
      construct a new {\tt Tweet} and store it inside of the {\tt ArrayList} for the valid instances of the {\tt Tweet}
      class.  If the message is invalid, then it should not construct a {\tt Tweet} and instead store the {\tt String}
      message in the {\tt ArrayList} for invalid tweet messages.

    \item Outputs all of the {\tt Tweet}s stored in the {\tt ArrayList} for the valid tweet messages. 

    \item Outputs all of the {\tt Tweet}s stored in the {\tt ArrayList} for the invalid tweet messages. 

  \end{enumerate}

  In the next laboratory assignment we will finish implementing a {\tt GatorTweet} program that can interact with
  Twitter. For instance, we will add features that can authenticate a user with the Twitter servers, display a user's
  Twitter timeline, and post a {\tt Tweet} to the Twitter site.

\section*{Summary of the Required Deliverables}

  This assignment invites you to submit one printed version of the following deliverables: 

  \begin{enumerate} 
    \item A description of the meaning and purpose of the provided {\tt build.xml} file.
    
    \item Screenshots demonstrating that you could correctly run the {\tt build.xml} file.
    
    \item The final version of your {\tt Tweet.java} and {\tt GatorTweet.java} files.
    
    \item The output associated with running {\tt GatorTweet} with a sample ``Tweets.txt'' file. 
  
  \end{enumerate}

  Along with turning in a printed version of these deliverables, you should ensure that everything is also available in
  the repository that is named according to the convention {\tt cs112S2015-<your user name>}. Please note that students
  in the class are responsible for completing and submitting their own version of this assignment.    While it is
  acceptable for members of this class to have high-level conversations, you should not share source code or full
  command lines with your classmates.  Please see the instructor if you have questions about the policies for this
  laboratory assignment.

  \end{document}
