\documentclass[12pt]{article}             
\textwidth = 6.5in
\textheight = 9.05in
\topmargin 0.0in
\oddsidemargin 0.0in
\evensidemargin 0.0in

% set it so that subsubsections have numbers and they
% are displayed in the TOC (maybe hard to read, might want to disable)

\setcounter{secnumdepth}{3}
\setcounter{tocdepth}{3}

% define widow protection 
        
\def\widow#1{\vskip #1\vbadness10000\penalty-200\vskip-#1}

% define a little section heading that doesn't go with any number

\def\littlesection#1{
\widow{2cm}
\vskip 0.5cm
\noindent{\bf #1}
\vskip 0.1cm
\noindent
}

% A paraphrase mode that makes it easy to see the stuff that shouldn't
% stay in for the final proposal

\newdimen\tmpdim
\long\def\paraphrase#1{{\parskip=0pt\hfil\break
\tmpdim=\hsize\advance\tmpdim by -15pt\noindent%
\hbox to \hsize
{\vrule\hskip 3pt\vrule\hfil\hbox to \tmpdim{\vbox{\hsize=\tmpdim
\def\par{\leavevmode\endgraf}
\obeyspaces \obeylines 
\let\par=\endgraf
\bf #1}}}}}

\renewcommand{\baselinestretch}{1.2}    % must go before the begin of doc
\newtheorem{principle}{Principle}
\newtheorem{definition}{Definition}
\newtheorem{define}{Definition}
% go with the way that CC sets the margins

\begin{document}

\vspace*{3in}

\begin{center}

CS 580: Junior Seminar \\
%Wednesday, February 28, 2007 \\ \mbox{} \\
Pledge: \\

\end{center}

\newpage
                                   
{\bf Question 1} {\Large $\;\;\;\frac{}{10}$ }

\vspace*{.5in}

The assigned papers described different methods for regression testing
and automated fault localization.  Using an appropriately labeled
diagram and a clear listing of steps, explain the input(s), output(s),
and operation of one of the proposed techniques.

\newpage

{\bf Question 2} {\Large $\;\;\; \frac{}{10}$ }

\vspace*{.5in}

Several of the papers describe the results from conducting an
empirical study.  Why did the authors decide to conduct experiments in
an attempt to evaluate their testing techniques?  How do you
empirically assess a method for regression testing or fault
localization?

\newpage

{\bf Question 3} {\Large $\;\;\; \frac{}{10}$ }

\vspace*{.5in}

The assigned papers use tables and graphs to highlight the results of
several experimental studies.  Identify and explain at least two
interesting empirical trends that were discussed in these papers.
Whenever possible, your response to this question should include
source code segments, diagrams, data points, and graphs.

%% \newpage

%% {\bf Question 4} {\Large $\;\;\; \frac{}{10}$ }

%% \newpage

%% {\bf Question 5} {\Large $\;\;\; \frac{}{10}$ }

\end{document}
