%!TEX root=cs112S2015-lab10.tex
% mainfile: cs112S2015-lab10.tex

\input{labspre.tex}

\usepackage[compact]{titlesec}

\begin{document} \MYTITLE{Laboratory Assignment Ten: Performance Evaluation of Java Collections}
\MYHEADERS{Laboratory Assignment Ten}{Due: April 9, 2015}

\section*{Introduction}

The Java Collections Framework provides an implementation of a wide variety of data containers, such as the {\tt
ArrayList}, {\tt LinkedList}, and the {\tt Vector}. Each of these different containers are implemented in a different
way and thus exhibit different trade-offs in their performance. For instance, the {\tt ArrayList} is ``backed'' by an
array while the {\tt LinkedList} uses nodes that are connected together by references. How do you think that these
different implementation choices will influence their performance in different usage scenarios? To get started on this
assignment, you should review the content in Sections 7.1 through 7.5 of your textbook, with a focus on the content in
Section 7.5. To learn more about the classes provided by the Java Collections Framework, you should also read the
JavaDoc documentation of all the relevant classes and find online sites that explain their performance trade-offs. What
ideas do you have for evaluating the performance of these containers?

\section*{Using the Benchmarking Framework}

To start this laboratory assignment, you should return to the {\tt cs112S2015-share} Git repository and type the ``{\tt
git pull}'' command in the terminal window.  Now, you should have a {\tt lab10/} directory that you can explore further.
Once again, please make sure that you can find the source code in this new directory and you understand why the
directories in the assignment are structured the way that they are. Next, you should use GVim to study the source code
in the {\tt build.xml} file.  As in the past assignments, when editing a Java program you can type ``{\tt
ant compile}'' in your terminal and it will compile all of the Java classes and save the bytecode in the correct
subdirectories inside of the {\tt bin/} directory.  Please see the instructor if you cannot get this to work.

One aspect of a container's performance is how efficiently it supports adding data to the front of the list after the
list has already been populated with data. After you have used the build system to compile the {\tt ArrayListBenchmarks}
and the {\tt LinkedListBenchmarks} classes, you can run them in the terminal window through the build system. What
command do you have to type to run these benchmarks? How did you learn which command to type? These two benchmarks
should output a few lines of empirical results. What is the meaning of this output? Once you understand the output
better, please carefully run the benchmarks multiple times, recording data values and noting important trends as you
continue. Why are these performance results evident in your output?

\section*{Adding New Performance Evaluation Benchmarks}

If you review the source code and the build system, you will notice that a benchmark for the {\tt Vector} class is
currently missing. As such, your next task for this assignment is to create a new class, called {\tt VectorBenchmarks}
that follows the same format as the two provided benchmarks. Additionally, you should add a new rule to the {\tt
build.xml} file so that you can easily run the benchmark by typing ``{\tt ant VectorBenchmarks}'' in the terminal
window. Does your new benchmark work correctly?

Of course, adding data to the front of an already populated container is only one operation that is likely to occur in
practice. For the next phase of this assignment, you should again review your notes about the Java collections framework
and conceptually design a completely new benchmark. Your benchmark should have a clear purpose and should be added, as
one or more methods, to the existing Java classes for the first benchmark. As you are implementing your new benchmark,
please make sure that you have a way to evaluate performance for all three of the Java containers that we are studying
in this laboratory assignment (e.g., the {\tt ArrayList}, {\tt LinkedList}, and {\tt Vector}). Now, you should repeat
this entire process again so that your final benchmarking framework contains three separate benchmarks. Please see the
course instructor if you have questions about this process.

\section*{Experimentally Evaluating the Efficiency of Java Collections}

Once you have finished implementing your benchmarks and testing them to ensure that they work correctly, you must plan
out and conduct a series of experiments to evaluate the efficiency of the Java Collections Framework. One aspect on this
experimentation effort is identifying all of the parameters of your benchmarks. For instance, the {\tt
ArrayListBenchmarks} class has two configurable parameters: {\tt initialListSize} and {\tt addToFrontNumber}. After you
have listed all of the parameters, you should determine what values you will assign to them, perhaps following the
doubling strategy to effectively pick your values. Now, go ahead and conduct all of your experiments, producing tables
of data and graphs to highlight your results. Finally, you should write a detailed experiment report that explains the
purpose of each benchmark, your hypothesis for the results, a justification for each stated hypothesis, the analysis of
your empirical results, and suggestions for future work.

\section*{Summary of the Required Deliverables}

  This assignment invites you to submit a signed and printed version of the following deliverables:

  \begin{enumerate}
  \itemsep0pt

  \item A complete description of the features provided by the Java Collections Framework.

  \item The properly commented version of every benchmark that you have implemented.

  \item The properly commented version of the build system that runs every benchmark.

  \item The output from several separate runs of your benchmarks, demonstrating all key features.

  \item A written report that provides a response to all of the questions posed in this assignment.

  \item A written report that follows all of the guidelines outlined in the previous section.

  \item A reflective commentary on the challenges that you faced when implementing the benchmarks.

  \item A reflective commentary on the challenges that you faced when doing the empirical analysis.

  \end{enumerate}

  Along with turning in a printed version of these deliverables, you should ensure that everything is also available in
  the repository that is named according to the convention {\tt cs112S2015-<your user name>}. Please note that students
  in the class are responsible for completing and submitting their own version of this assignment.    While it is
  acceptable for members of this class to have high-level conversations, you should not share source code or full
  command lines with your classmates.  Deliverables that are nearly identical to the work of others will be taken as
  evidence of violating the \mbox{Honor Code}.  Please see the instructor if you have questions about the policies for
  this assignment.

  \end{document}
